Surface programmable materials

### 2024-08-17 paper/main.tex
udpate

An elementary cellular automaton is a 1-dimensional cellular
automaton\footnote{https://en.wikipedia.org/wiki/Elementary_cellular_automaton}
where there are two possible states (labeled by 0 and 1). The rule to
determine the state of the cell in next generation depends only on the
current state of the cell and its two immediate neighbors.

@@

Rule 110 has been shown to be Turing Complete\footnote{Cook, M. (2009).
  A Concrete View of Rule 110 Computation. Electronic Proceedings in
  Theoretical Computer Science, 1, 31--55.
  https://doi.org/10.4204/EPTCS.1.4}, and thus capable of universal
computation.

@@

We can encode the Rule 110 cellular automaton into a Weighted Maximum
Independent Set Problem, with blue vertices assigned a weight of 1 and
red vertices assigned a weight of 2, as follows.
\includegraphics[width=3.125in,height=\textheight,keepaspectratio]{../notes/images/image.png}

This graph can be embedded into a grid graph, where two vertices are
connected if and only if their Euclidean distance is no more than $2$.
\includegraphics[width=3.125in,height=\textheight,keepaspectratio]{../notes/images/image-1.png}

@@

\begin{figure}
\centering
\includegraphics[width=3.125in,height=\textheight,keepaspectratio]{../notes/images/rule110.png}
\caption{Alt text}
\end{figure}

There are exactly \textbf{8} different MWIS solutions in this graph (the
weighted size of each MWIS solution is 7), each corresponding to one of
the \textbf{8} possible outputs of the automaton. We list them as
follows.
\includegraphics[width=5.20833in,height=\textheight,keepaspectratio]{../notes/images/gadget110.png}

@@

The gadget we constructed based on the Rule 110 cellular automaton
naturally possesses Turing completeness. Therefore, it can be tiled in a
two-dimensional plane to create a computational material with logical
operation capabilities. Utilizing copy gadget and cross
gadget\footnote{Nguyen, M.-T., Liu, J.-G., Wurtz, J., Lukin, M. D.,
  Wang, S.-T., \& Pichler, H. (2023). Quantum Optimization with
  Arbitrary Connectivity Using Rydberg Atom Arrays. PRX Quantum, 4(1),
  010316. https://doi.org/10.1103/PRXQuantum.4.010316}, we construct a
\textbf{Surface Programmable Material} with open boundary conditions as
follows.

@@

\begin{figure}
\centering
\includegraphics[width=3.125in,height=\textheight,keepaspectratio]{../notes/images/rule110_2-2_automaton.png}
\caption{Alt text}
\end{figure}

@@

The trade-off between the energy consumption and the speed of
computation~\cite{Feynman2018}. To avoid confusion, we emphasize the
``energy consumption'' is defined as the work done in a computational
process, which is the same as the amount of heat dissipated to the
environment. This quantity has a lower bound given by the Landauer
principle, which states that the work done in a computation is at least
$kT\ln 2$ per bit erased\footnote{Reeb, D. \& Wolf, M. M. An improved
  Landauer principle with finite-size corrections. New Journal of
  Physics 16, 1--34 (2014).}.

@@

Solving the ground state of the Hamiltonian of the Surface Programmable
Material is at least as hard as solving the circuit satisfiability
problem, which is NP-complete\footnote{Moore, Cristopher, and Stephan
  Mertens. The nature of computation. OUP Oxford, 2011.}.

@@

If the time evolution is sufficiently slow, then by the adiabatic
theorem, the system follows the instantaneous ground state, ending up in
the solution to the MWIS problem\footnote{Pichler, H., Wang, S.-T.,
  Zhou, L., Choi, S., \& Lukin, M. D. (2018). Quantum Optimization for
  Maximum Independent Set Using Rydberg Atom Arrays (arXiv:1808.10816).
  arXiv. http://arxiv.org/abs/1808.10816}.Then we only need to evalute
the minimum energy gap $\Delta_{QAA}$ between the ground and
first-excited states of instantaneous hamiltonian.

@@

\begin{figure}
\centering
\includegraphics[width=5.20833in,height=\textheight,keepaspectratio]{../notes/images/energy_gap_1_gadget.png}
\caption{Alt text}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[keepaspectratio]{../notes/images/error vs runtime.png}
\caption{Alt text}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[keepaspectratio]{../notes/images/toy_model_gradient_time_vs_depth_1_3.png}
\caption{Alt text}
\end{figure}

@@

Inspired by the way clothes are ironed, we conceived the idea of slowly
scanning a wave-shaped temperature curve from one end to the other. This
method is widely used in studying the heating-recrystallization
properties of real materials\footnote{Zhang, X., Yager, K. G., Douglas,
  J. F., \& Karim, A. (2014). Suppression of target patterns in domain
  aligned cold-zone annealed block copolymer films with immobilized
  film-spanning nanoparticles. Soft matter, 10(20), 3656-3666.}.

\begin{figure}
\centering
\includegraphics[keepaspectratio]{../notes/images/zhang2014.png}
\caption{Alt text}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[keepaspectratio]{../notes/images/toy_model_wave_packet_time_vs_depth_1_3.png}
\caption{Alt text}
\end{figure}

@@

The classical simulated annealing can only deal with the situation that
the temperature is space-invariant. Although there are many ways to
simulated a real-world system with temperature gradient, such as
Molecular Dynamics\footnote{Bai, X. M., Zhang, Y., \& Tonks, M. R.
  (2015). Testing thermal gradient driving force for grain boundary
  migration using molecular dynamics simulations. Acta materialia, 85,
  95-106.} \footnote{Deng, D., \& Murakawa, H. (2006). Numerical
  simulation of temperature field and residual stress in multi-pass
  welds in stainless steel pipe and comparison with experimental
  measurements. Computational materials science, 37(3), 269-277.}, but
these methods mainly depend on dynamically adjusting the atomic
velocities to achieve a temperature gradient, which can't be directly
used in our model.

There are also Monte Carlo-based methods to simulate a moving
temperature wave packet that facilitates material growth\footnote{Godfrey,
  A. W., \& Martin, J. W. (1995). Some Monte Carlo studies of grain
  growth in a temperature gradient. Philosophical Magazine A, 72(3),
  737-749.} \footnote{Tan, Y., Maniatty, A. M., Zheng, C., \& Wen, J. T.
  (2017). Monte Carlo grain growth modeling with local temperature
  gradients. Modelling and Simulation in Materials Science and
  Engineering, 25(6), 065003.}. However, these methods fail to establish
a direct connection between the simulated temperature and experimental
temperature\footnote{Zöllner, D. (2014). A new point of view to
  determine the simulation temperature for the Potts model simulation of
  grain growth. Computational materials science, 86, 99-107.}.Therefore,
based on HeatBath acceptance rule, we made a slight modification to the
classical SA model.

@@

\[
P = \frac{1}{1 + e^{\sum_i \frac{\Delta E_i}{T_i}}}
\]

@@

To examine whether this model could achieve a better performance when
using adiabatic methods, we need to generalize this energy function to
some good hamiltonian. There has already been work on mapping
combinatorial optimization problems to a spin-glass model (known as the
Ising Machine) and leveraging the properties of digital devices that
simulate this Ising Machine to find an appropriate ground state, which
encodes computational information\footnote{Aadit, N. A., Grimaldi, A.,
  Carpentieri, M., Theogarajan, L., Martinis, J. M., Finocchio, G., \&
  Camsari, K. Y. (2022). Massively parallel probabilistic computing with
  sparse Ising machines. Nature Electronics, 5(7), 460-468.}
\footnote{Bybee, C., Kleyko, D., Nikonov, D. E., Khosrowshahi, A.,
  Olshausen, B. A., \& Sommer, F. T. (2023). Efficient optimization with
  higher-order Ising machines. Nature Communications, 14(1), 6033.}.

@@

\[
H = \sum_{u,v \in E} J_{u,v}s_us_v + \sum_{i\in V}h_i s_i
\]

@@

\[
\min_{x \in \mathbb{R}^n} \sum_{i=1}^n c_ix_i\\
\text{s.t. } l_j \leq \sum_{i=1}^n a_{i,j}x_i \leq u_j, j=1...m\\
p_i \leq x_i \leq q_i, i=1...n
\]

@@

\begin{align*}
&\min_{J \in \mathbb{R}^{}{N(N{-}1)/2}, h\in \mathbb{R}^{}N} 0
&H(\mathbf s_i) \textless{} H(\mathbf s_j), \forall \mathbf s_i \in S_{\text{min}}, \mathbf s_j \in S \setminus S_{\text{min}}
&H(\mathbf s_i) = H(\mathbf s_j), \forall \mathbf s_i, \mathbf s_j \in S_{\text{min}}
\end{align*}

@@

\begin{equation}
J = \begin{pmatrix}
\cdot \& 1 \& 1 \& 2 \& 3
\cdot \& \cdot \& 2 \& 2 \& 5
\cdot \& \cdot \& \cdot \& 2 \& 5
\cdot \& \cdot \& \cdot \& \cdot \& 6
\cdot \& \cdot \& \cdot \& \cdot \& \cdot
\end{pmatrix}, h = \begin{pmatrix}
1
2
2
2
5
\end{pmatrix}
\end{equation}

@@

\subsection{References}\label{references}

\end{document}
\begin{acknowledgments}
    We thank Lei Wang, Madelyn Cain and XXX for helpful discussions on the simulation methods.
\end{acknowledgments}

@@

\begin{equation}
    H_{\text{Ryd}} = \sum_v \dfrac{\Omega_v}{2} \sigma^x_v -\sum_v \Delta_v n_v + \sum_{v \textless{} w}  V_{\text{Ryd}}(|\overrightarrow{\mathbf{r}_v} -\overrightarrow{\mathbf{r}_w}|)n_v n_w.
\end{equation}

@@

The NOR gate can be implemented using the Rydberg Hamiltonian (subfigure
c below). The NOR gate is a universal gate for classical computation.
\includegraphics[width=3.125in,height=\textheight,keepaspectratio]{../notes/images/gadgets.png}

@@

\textbf{Statement 4}: The Rydberg Hamiltonian, if cooled successfully
with some vertices fixed to certain configuration, can be used to solve
the circuit satisfiability problem, which is NP-complete.\footnote{Moore,
  Cristopher, and Stephan Mertens. The nature of computation. OUP
  Oxford, 2011.}

>>>>

An elementary cellular automaton is a 1-dimensional cellular
automaton, where there are two possible states (labeled by 0 and 1). The rule to
determine the state of the cell in next generation depends only on the
current state of the cell and its two immediate neighbors.

@@

Rule 110 has been shown to be Turing Complete~\cite{Cook2009}, and thus capable of universal
computation.

@@

We can encode the Rule 110 cellular automaton into a Weighted Maximum
Independent Set Problem, with blue vertices assigned a weight of 1 and
red vertices assigned a weight of 2, as follows.
\includegraphics[width=\columnwidth]{../notes/images/image.png}

This graph can be embedded into a grid graph, where two vertices are
connected if and only if their Euclidean distance is no more than $2$.
\includegraphics[width=\columnwidth]{../notes/images/image-1.png}

@@

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{../notes/images/rule110.png}
\caption{Alt text}
\end{figure}

There are exactly \textbf{8} different MWIS solutions in this graph (the
weighted size of each MWIS solution is 7), each corresponding to one of
the \textbf{8} possible outputs of the automaton. We list them as
follows.
\includegraphics[width=\columnwidth]{../notes/images/gadget110.png}

@@

The gadget we constructed based on the Rule 110 cellular automaton
naturally possesses Turing completeness. Therefore, it can be tiled in a
two-dimensional plane to create a computational material with logical
operation capabilities. Utilizing copy gadget and cross
gadget~\cite{Nguyen2023}, we construct a
\textbf{Surface Programmable Material} with open boundary conditions as
follows.

@@

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{../notes/images/rule110_2-2_automaton.png}
\caption{Alt text}
\end{figure}

@@

The trade-off between the energy consumption and the speed of
computation~\cite{Feynman2018}. To avoid confusion, we emphasize the
``energy consumption'' is defined as the work done in a computational
process, which is the same as the amount of heat dissipated to the
environment. This quantity has a lower bound given by the Landauer
principle, which states that the work done in a computation is at least
$kT\ln 2$ per bit erased\cite{Reeb2014}.

@@

Solving the ground state of the Hamiltonian of the Surface Programmable
Material is at least as hard as solving the circuit satisfiability
problem, which is NP-complete~\cite{Moore2011}.

@@

If the time evolution is sufficiently slow, then by the adiabatic
theorem, the system follows the instantaneous ground state, ending up in
the solution to the MWIS problem~\cite{Pichler2018}.
Then we only need to evalute
the minimum energy gap $\Delta_{QAA}$ between the ground and
first-excited states of instantaneous hamiltonian.

@@

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{../notes/images/energy_gap_1_gadget.png}
\caption{Alt text}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{../notes/images/error vs runtime.png}
\caption{Alt text}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{../notes/images/toy_model_gradient_time_vs_depth_1_3.png}
\caption{Alt text}
\end{figure}

@@

Inspired by the way clothes are ironed, we conceived the idea of slowly
scanning a wave-shaped temperature curve from one end to the other. This
method is widely used in studying the heating-recrystallization
properties of real materials~\cite{Zhang2014}.

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{../notes/images/zhang2014.png}
\caption{Alt text}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=\columnwidth]{../notes/images/toy_model_wave_packet_time_vs_depth_1_3.png}
\caption{Alt text}
\end{figure}

@@

The classical simulated annealing can only deal with the situation that
the temperature is space-invariant. Although there are many ways to
simulated a real-world system with temperature gradient, such as
Molecular Dynamics~\cite{Bai2015, Deng2006}, but
these methods mainly depend on dynamically adjusting the atomic
velocities to achieve a temperature gradient, which can't be directly
used in our model.

There are also Monte Carlo-based methods to simulate a moving
temperature wave packet that facilitates material growth~\cite{Godfrey1995, Tan2017}.
However, these methods fail to establish a direct connection between the simulated temperature and experimental
temperature~\cite{Zollner2014}.
Therefore, based on HeatBath acceptance rule, we made a slight modification to the classical SA model.

@@

\begin{equation}
P = \frac{1}{1 + e^{\sum_i \frac{\Delta E_i}{T_i}}}
\end{equation}

@@

To examine whether this model could achieve a better performance when
using adiabatic methods, we need to generalize this energy function to
some good hamiltonian. There has already been work on mapping
combinatorial optimization problems to a spin-glass model (known as the
Ising Machine) and leveraging the properties of digital devices that
simulate this Ising Machine to find an appropriate ground state, which
encodes computational information~\cite{Aadit2022,Bybee2023}.

@@

\begin{equation}
H = \sum_{u,v \in E} J_{u,v}s_us_v + \sum_{i\in V}h_i s_i
\end{equation}

@@

\begin{equation}
    \begin{split}
        &\min_{x \in \mathbb{R}^n} \sum_{i=1}^n c_ix_i\\
        &\text{s.t. } l_j \leq \sum_{i=1}^n a_{i,j}x_i \leq u_j, \; j=1,\ldots,m\\
        &p_i \leq x_i \leq q_i, \; i=1,\ldots, n
    \end{split}
\end{equation}

@@

\begin{equation}
    \begin{split}
        &\min_{J \in \mathbb{R}^{N(N{-}1)/2}, h\in \mathbb{R}^N} 0\\
        &H(\mathbf s_i) < H(\mathbf s_j), \forall \mathbf s_i \in S_{\text{min}}, \mathbf s_j \in S \setminus S_{\text{min}}\\
        &H(\mathbf s_i) = H(\mathbf s_j), \forall \mathbf s_i, \mathbf s_j \in S_{\text{min}}
    \end{split}
\end{equation}

@@

\begin{equation}
J = \begin{pmatrix}
\cdot & 1 & 1 & 2 & 3\\
\cdot & \cdot & 2 & 2 & 5\\
\cdot & \cdot & \cdot & 2 & 5\\
\cdot & \cdot & \cdot & \cdot & 6\\
\cdot & \cdot & \cdot & \cdot & \cdot
\end{pmatrix}, h = \begin{pmatrix}
1\\
2\\
2\\
2\\
5
\end{pmatrix}
\end{equation}

@@

\begin{acknowledgments}
    We thank Lei Wang, Madelyn Cain and XXX for helpful discussions on the simulation methods.
\end{acknowledgments}

@@

\begin{equation}
    H_{\text{Ryd}} = \sum_v \dfrac{\Omega_v}{2} \sigma^x_v -\sum_v \Delta_v n_v + \sum_{v < w}  V_{\text{Ryd}}(|\overrightarrow{\mathbf{r}_v} -\overrightarrow{\mathbf{r}_w}|)n_v n_w.
\end{equation}

@@

The NOR gate can be implemented using the Rydberg Hamiltonian (subfigure
c below). The NOR gate is a universal gate for classical computation.
\includegraphics[width=\columnwidth]{../notes/images/gadgets.png}

@@

\textbf{Statement 4}: The Rydberg Hamiltonian, if cooled successfully
with some vertices fixed to certain configuration, can be used to solve
the circuit satisfiability problem, which is NP-complete~\cite{Moore2011}.
